\begin{abstract}
Contrastive learning applied to self-supervised representation learning has seen a resurgence in recent years, leading to state of the art performance in the unsupervised training of deep image models. Modern batch contrastive approaches subsume or significantly outperform traditional contrastive losses such as triplet, max-margin and the N-pairs loss. In this work, we extend the self-supervised batch contrastive approach to the \emph{fully-supervised} setting, allowing us to effectively leverage label information. Clusters of points belonging to the same class are pulled together in embedding space, while simultaneously pushing apart clusters of samples from different classes. We analyze two possible versions of the supervised contrastive (SupCon) loss, identifying the best-performing formulation of the loss. On ResNet-200, we achieve top-1 accuracy of $81.4\%$ on the ImageNet dataset, which is $0.8\%$ above the best number reported for this architecture. We show consistent outperformance over cross-entropy on other datasets and two ResNet variants. The loss shows benefits for robustness to natural corruptions, and is more stable to hyperparameter settings such as optimizers and data augmentations. %In reduced data settings, it outperforms cross-entropy significantly.
Our loss function is simple to implement and reference TensorFlow code is released at \url{https://t.ly/supcon} \footnote[1]{\scriptsize PyTorch implementation: \url{https://github.com/HobbitLong/SupContrast}}.


\end{abstract}
